Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die Einhaltung der Echtzeit

### 2024-06-11 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] Host Configurations finished

\begin{table}[h]
	\centering
	\caption[System configuration]{System configuration}
	\label{tab:testbed_configuration}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
	\begin{tabular}{|l|l|}
	\hline
	\textbf{CPU} & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
	\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB \\ \hline
	\textbf{GPU} & NVIDIA RTX A500 Laptop GPU \\ \hline
	\textbf{BIOS} & Dell Version 1.12.0 \\ \hline
	\textbf{OS} & Ubuntu 22.04.4 LTS \\
	\hline
	\end{tabular}}
	\end{table}

@@

\begin{table}[h]
	\centering
	\caption{Latency statistics of Salamander 4 bare metal in microseconds}
	\label{tab:latency_statistics_bare_metal}
	\begin{tabular}{|c|c|}\hline
	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
	Average Latency & 4.06 \\\hline
	Maximum Latency & 10.71 \\\hline
	Minimum Latency & 2.82 \\\hline
	Standard Deviation & 0.85 \\\hline
	\end{tabular}
	\end{table}

@@

\begin{table}[h]
	\centering
	\caption{Latency statistics of default Salamander 4 virtualization in microseconds}
	\label{tab:latency_statistics_virt}
	\begin{tabular}{|c|c|}\hline
	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
	Average Latency & 46.22 \\\hline
	Maximum Latency & 707.62 \\\hline
	Minimum Latency & 15.59 \\\hline
	Standard Deviation & 52.13 \\\hline
	\end{tabular}
	\end{table}

@@

\begin{table}[h]
	\centering
	\caption{BIOS Configurations}
	\label{tab:bios_configuration}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
	\begin{tabular}{|l|l|}
	\hline
	\textbf{Option} & \textbf{Status} \\
	\hline
	Hyper Threading & Disabled \\
	\hline
	Intel SpeedStep® & Disabled \\
	\hline
	Intel® Speed Shift Technology & Disabled \\
	\hline
	C States & Disabled \\
	\hline
	VT-d & Enabled \\
	\hline
	\end{tabular}}
	\end{table}

@@

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
	\begin{lstlisting}[name={User and Kernel Tasks},label={output:pse},escapeinside={(*@}{@*)}]
		sigma_ibo@sigma-ibo:~$ cat /sys/devices/system/cpu/isolated
		4
		sigma_ibo@sigma-ibo:~$ ps axHo psr,pid,lwp,args,policy,nice,rtprio | awk '$1 == 4'
		4      38      38 [cpuhp/4]                   TS    0      -
		4      39      39 [idle_inject/4]             FF    -     50
		4      40      40 [migration/4]               FF    -     99
		4      41      41 [ksoftirqd/4]               TS    0      -
		4      42      42 [kworker/4:0-events]        TS    0      -
		4      43      43 [kworker/4:0H-kblockd]      TS  -20      -
		4     153     153 [kworker/4:1-events]        TS    0      -
		4   81649   81649 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   81654 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   81676 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   81702 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   82185 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   82187 qemu-system-x86_64 -M pc,ac TS    0      -
		4   82134   82134 [kworker/4:1H-kblockd]      TS  -20      -
\end{lstlisting}
\end{minipage}
\subsection{Interrupt Affinity}\label{subsec:irq_handling}
Once the CPUs were isolated, interrupt requests handling was the next step. The purpose of interrupt requests is to inform the CPU to stop working on a certain job and start working on another. This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems. The ~\texttt{/proc/interrupts} file can be monitored using the command ~\texttt{watch -d -c cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time. The IRQs needed to be removed from the isolated CPU by manipulating the ~\texttt{/proc/irq/<IRQ>/smp\_affinity} files. The value in the ~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU. The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU13). When there are 14 CPUs available, the default value for ~\texttt{smp\_affinity} would be 3FFF. Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 3FEF (11 1111 1110 1111). The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU. By changing the values of the ~\texttt{smp\_affinity} files, the assignment of IRQs to CPUs is controlled. However, as the proc filesystem resets to its default state after each reboot, manually changing numerous IRQ files is tedious and time-consuming. This process was automated using the shell script in Code~\ref{script:change_smp_affinity}, which is executed after every reboot.

@@

		# Check if there were any successful IRQs
		if [ ${#succeeded_IRQs[@]} -ne 0 ]; then
			# Remove the first entry from the succeeded_IRQs array
			succeeded_IRQs=("${succeeded_IRQs[@]:1}")
			echo "IRQs ${succeeded_IRQs[@]} were removed from CPU $CPU."
		fi		
\end{lstlisting}
\end{minipage}
\subsection{Disable RT throttling}
If a real-time task consumes 100\% of the CPU time, the system may become unresponsive as a whole. This happens because an RT process is constantly using the CPU and the Linux scheduler will not schedule other non-RT processes in the meantime. To prevent complete system lockups, the kernel has a function to throttle RT processes if they consume 0.95 seconds out of every 1 second of CPU time. It does this by pausing the process for the remaining 0.05 seconds, which is not desired because this could result in missed deadlines. RT throttling can be disabled by  writing the value -1 to the ~\texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file. This change also needs to be made permanent because the proc filesystem resets to its default state after each reboot. For this purpose, the line ~\texttt{kernel.sched\_rt\_runtime\_us = -1} can be appended to the end of the ~\texttt{/etc/sysctl.conf} file, which is read at boot time and used to configure kernel parameters. This reduces the potential for missed deadlines. 
\subsection{Disable RCU CPU stall warnings}
As already mentioned in section \ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data. An RCU CPU stall in the Linux kernel can occur due to several reasons. These include a CPU looping in an RCU read-side critical section, a CPU looping with interrupts disabled, a CPU looping with preemption disabled, or a CPU not getting around to less urgent tasks, known as “bottom halves”. The number of seconds the kernel should wait before checking for stalled CPUs and reporting a stall warning can be set via the  ~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_timeout} file. These warnings can be suppressed alltogether to reduce the potential for increased latency by writing ``1'' to the ~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file. This setting is also not persistent across reboots, so the command needs to be added to a startup script.
\subsection{RT-priority} 
\subsection{No unexpected RT processes}
It is important that there are no other unexpected real time processes running on the system. A high priority RT process could occasionally be created by the host operating system at boot, hence disabling the offending RT process is advised if it is not needed. The process of disabling these RT processes is usually described in the documentation for the chosen Linux distribution.
\subsection{Set CPU Affinity of IRQ thread to CPU 0}

\subsection{Set Device Driver Work Queue to CPU 0}

\subsection{Stop Certain Services}

\subsection{Disable Machine Check}

There are other configurations that may be relevant depending on your use case, some of which are documented in [this talk](https://www.youtube.com/watch?v=NrjXEaTSyrw) and [this other talk](https://www.youtube.com/watch?v=w3yT8zJe0Uw). Additionally, quality-of-life configurations, such the variables in /etc/security/limits.conf, may need to be tuned as well.

``korrekt''

@@

\clearpage
\section{Guest OS Configurations}

@@

\chapter{To Include}\label{cha:to_include}
\section{Latency Comparison}
In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4. For this purpose, the latency tool of the Xenomai test suite was used. The latency was measured under two conditions, idle and CPU-stressed. The goal was to optimize the latency of the virtualization of Salamander 4 OS to closely match that of the bare metal version.

@@

\begin{table}[h!]
	\centering
	\begin{minipage}{.5\textwidth}
	\centering
	\caption{Host report CPU19 (Total of 445.908)}
	\label{tab:results_host_report}
	\begin{tabular}{|c|c|c|}
	\hline
	PID & Task & Count \\ \hline
	182579 & qemu-system-x86 & 302.748 \\ \hline
	0 & \textless{}idle\textgreater{} & 112.911 \\ \hline
	182618 & vhost & 21.204 \\ \hline
	182572 & qemu-system-x86 & 7.597 \\ \hline
	182755 & qemu-system-x86 & 644 \\ \hline
	182754 & qemu-system-x86 & 643 \\ \hline
	181870 & kworker/19:1 & 139 \\ \hline
	3820 & kworker/19:1H & 16 \\ \hline
	94 & migration/19 & 6 \\ \hline
	\end{tabular}
	\end{minipage}%
	\begin{minipage}{.5\textwidth}
	\centering
	\caption{Guest report (Total of 362.370)}
	\label{tab:results_guest_report}
	\begin{tabular}{|c|c|c|}
	\hline
	PID & Task & Count \\ \hline
	0 & \textless{}idle\textgreater{} & 150.744 \\ \hline
	331 & LRT-Main & 56.697 \\ \hline
	377 & trace-cmd & 48.507 \\ \hline
	346 & CLI & 26.426 \\ \hline
	378 & kthreadd & 25.837 \\ \hline
	340 & MainTaskLow & 19.291 \\ \hline
	339 & \textless{}...\textgreater{} & 9.980 \\ \hline
	34 & MainTaskHigh & 9.185 \\ \hline
	327 & LE-Logger & 4.965 \\ \hline
	369 & kworker/0:0 & 2.793 \\ \hline
	321 & kWorker-LRT & 2.542 \\ \hline
	328 & LRTMgr-Main & 1.651 \\ \hline
	34 & LrtMgrCyclic & 1.220 \\ \hline
	332 & cobalt\_printf & 1.112 \\ \hline
	325 & LE-System & 534 \\ \hline
	343 & TCP-Listen & 187 \\ \hline
	15 & rcu\_preempt & 162 \\ \hline
	25 & kcompactd0 & 122 \\ \hline
	58 & kworker/0:1H & 96 \\ \hline
	63 & kworker/u2:2 & 89 \\ \hline
	8 & jbd2/sda-8 & 86 \\ \hline
	22 & kworker/0:1 & 56 \\ \hline
	1 & init & 31 \\ \hline
	2 & kthreadd & 25 \\ \hline
	375 & trace-cmd & 24 \\ \hline
	14 & ksoftirqd/0 & 8 \\ \hline
	\end{tabular}
	\end{minipage}
	\end{table}

@@

\begin{table}[h]
	\centering
	\begin{tabular}{|p{0.1\textwidth}|p{0.45\textwidth}|p{0.45\textwidth}|}
	\hline
	\textbf{Number} & \textbf{Book} & \textbf{Citation} \\
	\hline
	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core. \\
	\hline
	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core. \\
	\hline
	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core. \\
	\hline
	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core. \\
	\hline
	\end{tabular}
	\caption{Your Caption}
	\label{tab:my_label}
	\end{table}

>>>>

\begin{table}[H]
	\centering
	\caption[System configuration]{System configuration}
	\label{tab:testbed_configuration}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
	\begin{tabular}{|l|l|}
	\hline
	\textbf{CPU} & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
	\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB \\ \hline
	\textbf{GPU} & NVIDIA RTX A500 Laptop GPU \\ \hline
	\textbf{BIOS} & Dell Version 1.12.0 \\ \hline
	\textbf{OS} & Ubuntu 22.04.4 LTS \\
	\hline
	\end{tabular}}
	\end{table}

@@

\begin{table}[H]
	\centering
	\caption{Latency statistics of Salamander 4 bare metal in microseconds}
	\label{tab:latency_statistics_bare_metal}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
	\begin{tabular}{|c|c|}\hline
	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
	Average Latency & 4.06 \\\hline
	Maximum Latency & 10.71 \\\hline
	Minimum Latency & 2.82 \\\hline
	Standard Deviation & 0.85 \\\hline
	\end{tabular}}
	\end{table}

@@

\begin{table}[H]
	\centering
	\caption{Latency statistics of default Salamander 4 virtualization in microseconds}
	\label{tab:latency_statistics_virt}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
	\begin{tabular}{|c|c|}\hline
	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
	Average Latency & 46.22 \\\hline
	Maximum Latency & 707.62 \\\hline
	Minimum Latency & 15.59 \\\hline
	Standard Deviation & 52.13 \\\hline
	\end{tabular}}
	\end{table}

@@

\begin{table}[H]
	\centering
	\caption{BIOS Configurations}
	\label{tab:bios_configuration}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
	\begin{tabular}{|l|l|}
	\hline
	\textbf{Option} & \textbf{Status} \\
	\hline
	Hyper Threading & Disabled \\
	\hline
	Intel SpeedStep® & Disabled \\
	\hline
	Intel® Speed Shift Technology & Disabled \\
	\hline
	C States & Disabled \\
	\hline
	VT-d & Enabled \\
	\hline
	\end{tabular}}
	\end{table}

@@

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
	\begin{lstlisting}[name={User and Kernel Tasks},label={output:pse},escapeinside={(*@}{@*)}]
		sigma_ibo@sigma-ibo:~$ cat /sys/devices/system/cpu/isolated
		4
		sigma_ibo@sigma-ibo:~$ ps axHo psr,pid,lwp,args,policy,nice,rtprio | awk '$1 == 4'
		4      38      38 [cpuhp/4]                   TS    0      -
		4      39      39 [idle_inject/4]             FF    -     50
		4      40      40 [migration/4]               FF    -     99
		4      41      41 [ksoftirqd/4]               TS    0      -
		4      42      42 [kworker/4:0-events]        TS    0      -
		4      43      43 [kworker/4:0H-kblockd]      TS  -20      -
		4     153     153 [kworker/4:1-events]        TS    0      -
		4   81649   81649 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   81654 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   81676 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   81702 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   82185 qemu-system-x86_64 -M pc,ac TS    0      -
		4   81649   82187 qemu-system-x86_64 -M pc,ac TS    0      -
		4   82134   82134 [kworker/4:1H-kblockd]      TS  -20      -
\end{lstlisting}
\end{minipage}
\subsection{Interrupt Affinity}\label{subsec:irq_handling}
Once the CPUs were isolated, interrupt requests handling was the next step. The purpose of interrupt requests is to inform the CPU to stop working on a certain job and start working on another. This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems. The ~\texttt{/proc/interrupts} file can be monitored using the command ~\texttt{watch -d -c cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time. The IRQs needed to be removed from the isolated CPU by manipulating the ~\texttt{/proc/irq/<IRQ>/smp\_affinity} files. The value in the ~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU. The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU13). When there are 14 CPUs available, the default value for ~\texttt{smp\_affinity} would be 3FFF. Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 3FEF (11 1111 1110 1111). The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU. By changing the values of the ~\texttt{smp\_affinity} files, the assignment of IRQs to CPUs is controlled. However, as the proc filesystem resets to its default state after each reboot, manually changing numerous IRQ files is tedious and time-consuming. This process was automated using the shell script in Code~\ref{script:change_smp_affinity}, which is executed after every reboot. Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads. This can be done through finding out the respective process IDs through \texttt{ps -e | grep 'irq/.*nvme'} and then executing \texttt{sudo taskset -a -p -c 0 <PID>}.

@@

		# Check if there were any successful IRQs
		if [ ${#succeeded_IRQs[@]} -ne 0 ]; then
			# Remove the first entry from the succeeded_IRQs array
			succeeded_IRQs=("${succeeded_IRQs[@]:1}")
			echo "IRQs ${succeeded_IRQs[@]} were removed from CPU $CPU."
		fi		
\end{lstlisting}
\end{minipage}
\subsection{Disable RT throttling}
If a real-time task consumes 100\% of the CPU time, the system may become unresponsive as a whole. This happens because an RT process is constantly using the CPU and the Linux scheduler will not schedule other non-RT processes in the meantime. To prevent complete system lockups, the kernel has a function to throttle RT processes if they consume 0.95 seconds out of every 1 second of CPU time. It does this by pausing the process for the remaining 0.05 seconds, which is not desired because this could result in missed deadlines. RT throttling can be disabled by  writing the value -1 to the ~\texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file. This change also needs to be made permanent because the proc filesystem resets to its default state after each reboot. For this purpose, the line ~\texttt{kernel.sched\_rt\_runtime\_us = -1} can be appended to the end of the ~\texttt{/etc/sysctl.conf} file, which is read at boot time and used to configure kernel parameters. This reduces the potential for missed deadlines. 
\subsection{Disable RCU CPU stall warnings}
As already mentioned in section \ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data. An RCU CPU stall in the Linux kernel can occur due to several reasons. These include a CPU looping in an RCU read-side critical section, a CPU looping with interrupts disabled, a CPU looping with preemption disabled, or a CPU not getting around to less urgent tasks, known as “bottom halves”. The number of seconds the kernel should wait before checking for stalled CPUs and reporting a stall warning can be set via the  ~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_timeout} file. These warnings can be suppressed alltogether to reduce the potential for increased latency by writing ``1'' to the ~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file. This setting is also not persistent across reboots, so the command needs to be added to a startup script.
\subsection{RT-priority} 
Having a real-time kernel itself is a crucial part of achieving deterministic behavior, but not enough to take full advantage of the real-time capabilities. One key aspect of this are real-time priorities, thoroughly explained by Richard Weinberger in~\cite{LinuxProcessPriorities}. In essence, Table \ref{tab:scheduling_priorities} lists minimum and maximum priorities for different scheduling policies in Linux.

\begin{table}[H]
	\centering
	\caption{Minimum and maximum priorities for different scheduling policies}
	\label{tab:scheduling_priorities}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
	\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Scheduling Policy} & \textbf{Min Priority} & \textbf{Max Priority} \\ \hline
	SCHED\_OTHER               & 0                     & 0                     \\ \hline
	SCHED\_FIFO                & 1                     & 99                    \\ \hline
	SCHED\_RR                  & 1                     & 99                    \\ \hline
	SCHED\_BATCH               & 0                     & 0                     \\ \hline
	SCHED\_IDLE                & 0                     & 0                     \\ \hline
	SCHED\_DEADLINE            & 0                     & 0                     \\ \hline
	\end{tabular}}
	\end{table}

\noindent \texttt{SCHED\_FIFO} allows deterministic, high-priority execution of critical tasks without being preempted by lower-priority processes. The virtual machine can either be started with \texttt{chrt -f <PRIO>} or adjusted at a later point with \texttt{chrt -f <PRIO> <PID>}.

\subsection{No unexpected RT processes}
It is important that there are no other unexpected real time processes running on the system. A high priority RT process could occasionally be created by the host operating system at boot, hence disabling the offending RT process is advised if it is not needed. The process of disabling these RT processes is usually described in the documentation for the chosen Linux distribution.
\subsection{Set Device Driver Work Queue}
The device driver work queue allows time-consuming tasks to be offloaded to be processed later in a separate kernel thread. By setting the work queue away from CPU 4 to another CPU, it is free to handle real-time tasks without being interrupted by these work queue tasks. This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask}.
\subsection{Stop Certain Services}
Services like irqbalance.service, thermald.service, and wpa\_supplicant.service, as presented in Table~\ref{table:stop_servies} can be further sources for random latency and unnecessary overhead. Stopping these services through \texttt{sudo systemctl stop <SERVICE>} means they will not be able to interrupt the CPU with their tasks.

\begin{table}[H]
	\centering
	\caption{Description of Services}
	\label{table:stop_servies}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
	\begin{tabular}{|l|l|}
	\hline
	\textbf{Service} & \textbf{Description} \\
	\hline
	irqbalance.service & Distributes hardware interrupts across CPUs \\\hline
	thermald.service & A daemon that prevents overheating \\\hline
	wpa\_supplicant.service & A service for wireless network devices \\
	\hline
	\end{tabular}}
	\end{table}
\subsection{Disable Machine Check}
Machine checks report hardware errors and these checks can cause interruptions and increase latency. Hence it is best to disable them in real-time scenarios by writing a ``0'' to the \texttt{/sys/devices/system/machinecheck/machinecheck0/check\_interval} file.

@@

\section{Guest OS Configurations}

\section{Other configurations}
There are other configurations that may be relevant depending on the case. Linux Kernel Developer Steven Rostedt explains all relevant aspects of a real-time system that must be considered in~\cite{kernelrecipesKernelRecipes20162016} and gives insight for finding sources of latency on the linux system in~\cite{thelinuxfoundationFindingSourcesLatency2020}. A Checklist for Writing Linux Real-Time Applications is provided by John Ogness in \cite{thelinuxfoundationChecklistWritingLinux2020}. Every layer of the system stack must be deterministic to ensure predictable and reliable latency, including hardware, operating system, middleware and drivers, and the application software. \cite{HOWTOBuildRTapplication} and \cite{RealtimeProgrammingLinux} describe the process of writing hard real time Linux programs using the real time preemption patch in great detail. Various hardware and software tunings are mentioned in~\cite{KVMQemuVirtualization} and~\cite{RealTimePerformanceTuning2022}.

@@

\chapter{To Include}\label{cha:to_include}
``korrekt''

\section{Latency Comparison}
In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4. For this purpose, the latency tool of the Xenomai test suite was used. The latency was measured under two conditions, idle and CPU-stressed. The goal was to optimize the latency of the virtualization of Salamander 4 OS to closely match that of the bare metal version.

@@

\begin{table}[H]
	\centering
	\begin{minipage}{.5\textwidth}
	\centering
	\caption{Host report CPU19 (Total of 445.908)}
	\label{tab:results_host_report}
	\begin{tabular}{|c|c|c|}
	\hline
	PID & Task & Count \\ \hline
	182579 & qemu-system-x86 & 302.748 \\ \hline
	0 & \textless{}idle\textgreater{} & 112.911 \\ \hline
	182618 & vhost & 21.204 \\ \hline
	182572 & qemu-system-x86 & 7.597 \\ \hline
	182755 & qemu-system-x86 & 644 \\ \hline
	182754 & qemu-system-x86 & 643 \\ \hline
	181870 & kworker/19:1 & 139 \\ \hline
	3820 & kworker/19:1H & 16 \\ \hline
	94 & migration/19 & 6 \\ \hline
	\end{tabular}
	\end{minipage}%
	\begin{minipage}{.5\textwidth}
	\centering
	\caption{Guest report (Total of 362.370)}
	\label{tab:results_guest_report}
	\begin{tabular}{|c|c|c|}
	\hline
	PID & Task & Count \\ \hline
	0 & \textless{}idle\textgreater{} & 150.744 \\ \hline
	331 & LRT-Main & 56.697 \\ \hline
	377 & trace-cmd & 48.507 \\ \hline
	346 & CLI & 26.426 \\ \hline
	378 & kthreadd & 25.837 \\ \hline
	340 & MainTaskLow & 19.291 \\ \hline
	339 & \textless{}...\textgreater{} & 9.980 \\ \hline
	34 & MainTaskHigh & 9.185 \\ \hline
	327 & LE-Logger & 4.965 \\ \hline
	369 & kworker/0:0 & 2.793 \\ \hline
	321 & kWorker-LRT & 2.542 \\ \hline
	328 & LRTMgr-Main & 1.651 \\ \hline
	34 & LrtMgrCyclic & 1.220 \\ \hline
	332 & cobalt\_printf & 1.112 \\ \hline
	325 & LE-System & 534 \\ \hline
	343 & TCP-Listen & 187 \\ \hline
	15 & rcu\_preempt & 162 \\ \hline
	25 & kcompactd0 & 122 \\ \hline
	58 & kworker/0:1H & 96 \\ \hline
	63 & kworker/u2:2 & 89 \\ \hline
	8 & jbd2/sda-8 & 86 \\ \hline
	22 & kworker/0:1 & 56 \\ \hline
	1 & init & 31 \\ \hline
	2 & kthreadd & 25 \\ \hline
	375 & trace-cmd & 24 \\ \hline
	14 & ksoftirqd/0 & 8 \\ \hline
	\end{tabular}
	\end{minipage}
	\end{table}

@@

\begin{table}[H]
	\centering
	\begin{tabular}{|p{0.1\textwidth}|p{0.45\textwidth}|p{0.45\textwidth}|}
	\hline
	\textbf{Number} & \textbf{Book} & \textbf{Citation} \\
	\hline
	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core. \\
	\hline
	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core. \\
	\hline
	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core. \\
	\hline
	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core. \\
	\hline
	\end{tabular}
	\caption{Your Caption}
	\label{tab:my_label}
	\end{table}
